---
title: "Student Demographics and Final Grades"
subtitle: "CIS 351 Final Project"
author:
    - name: Author 1
    - name: Author 2
    - name: Author 3
format: 
    html:
        embed-resources: true
toc: true
editor_options: 
  chunk_output_type: console
---

## Introduction

Say something (intelligent) about your dataset. Which variables are of interest to you and how they lead you to form the research questions.

You must have exactly two research questions.

## Loading packages and data

```{r}
#| message: false

library(tidyverse)

library(janitor)

library(ucimlrepo)

library(ggpubr)

library(car)

library(skimr)

library(gt)

uci_df <- fetch_ucirepo(name = 'Student Performance')

df <- uci_df$data$original |> 
    clean_names()
```

## Exploratory Data Analysis Using `skim()`

```{r}
skim(df)
```

## Research Questions

1. Is there a difference in the mean final scores between students from urban areas and those from rural areas?

2. Assume that `studytime` and the final score (`g3`) are continuous numeric variables. Are these two variables correlated?


## Examine variables of interest

### Question 1

The variables of interest are `address` and `g3`.

For `address` is a binary categorical variable. We can use the `summarize()` to create a table of counts by address type (i.e., `U` or `R`).

```{r}
df_rows <- nrow(df)

df |> 
    summarize(
        count = n(),
        percent = round(count/df_rows*100, 1),
        .by = address
    ) |> 
    gt()
```

Table shows that 197 students have rural addresses and 452 with urban addresses.

The second variable of interest is `g3`, which is a numeric variable. We can use the `summary()` function to create a five-number summary for `g3`.

```{r}
summary(df$g3) 
```

The mean and median are roughly equal, so it's likely that `g3` (final score) is symmetric.

We can create a two-way to look at the distributions of final scores (`g3`) by `address` using the `tabyl()` function.

```{r}
df |> 
    tabyl(address, g3) |> 
    gt()
```

### Question 2

The variable `g3` was discussed above, so we will proceed with the discussion of `studytime`.

```{r}
summary(df$studytime)
```
Note that the median and the mean for `studytime` are similar, but the histogram from `skim()` tells us that the distribution of the variable is not symmetric.

## Data visualizations

### Question 1

Since `g3` is a numeric variable, a histogram is an appropriate visualization.

```{r}
df |> 
    ggplot(
        aes(x = g3)
    ) +
    geom_histogram(
        fill = '#A90533',
        color = 'white'
    ) +
    theme_classic() +
    labs(
        title = 'Final Scores',
        x = 'Final Score (out of 20)',
        y = 'Count'
    )
```

To visualize the final scores by `address` type, we use the boxplot.


```{r}
df |> 
    ggplot(
        aes(
            x = address,
            y = g3,
            fill = address
        )
    ) +
    geom_boxplot() +
    theme_minimal() +
    labs(
        title = 'Final Scores by Address Type',
        x = 'Address Type',
        y = 'Final Scores'
    ) +
    scale_fill_manual(values = c('#A90533', '#FFB500')) #manually change the colors of the boxes; #A90533 and #FFB500 are the official primary colors of Flagler College
```
Observe that both groups have outliers. the distributions of the groups are not similar. 

### Question 2

Since `studytime` and `g3` are both continuous and numeric under the assumption made in the research questions section. The appropriate visualization for this question is the scatter plot.

```{r}
ggscatter(
    df,
    x = 'studytime',
    y = 'g3',
    add = 'reg.line',
    add.params = list(color = 'blue'),
    xlab = 'Study Time (Hours)',
    ylab = 'Final Score'
) +
    stat_cor(
        label.x = 1,
        label.y = 25
    ) +
    stat_regline_equation(
        label.x = 1,
        label.y = 22,
    )
```



## Statistical tests

### Question 1

Since sample sizes are sufficiently large, we can use the $t$-test. However, the variable type for final scores is discrete and numerical, we should ultimately use the rank-sum test.

$H_0$: $\mu_R = \mu_U$ <br>
$H_a$: $\mu_R \ne \mu_U$

```{r}
test_1 <- t.test(
    g3 ~ address,
    data = df
)

test_1
```
The $p$-value is $p = `{r} test_1$p.value` < 0.05$, so we reject the null hypothesis that the mean final scores are the same.

### Question 2

Note that `ggpubr` gave us the necessary statistical outputs for the regression analysis between `studytime` and final score (`g3`). We have the $p$-value of $p = 1.1 \times 10^{-10} < 0.05$. The correlation coefficient $R = 0.25$ with the regression equation $y = 10 + 0.97x$, where $x$ is `studytime` and $y$ is the final score (`g3`).

## Conclusion

Say something about your research questions and the results from your statistical tests.





